%%  $OpenXM: OpenXM/doc/OpenXM-specs/OpenXM.tex,v 1.10 2000/10/19 00:17:50 takayama Exp $
\documentclass{article}
\IfFileExists{epsfig.sty}{\usepackage{epsfig}}{}
\usepackage{html}
\begin{document}
/*&jp
\title{{\bf Open XM の設計と実装} \\
 --- Open message eXchange protocol for Mathematics }
\author{ 野呂正行, 
         高山信毅\thanks{神戸大学理学部数学教室,\htmladdnormallink{http://www.math.kobe-u.ac.jp/$\tilde{\ }$taka}{http://www.math.kobe-u.ac.jp/\tilde{\ }taka}}}
\date{ 2000年, 11月17日}
*/
/*&eg
\title{{\bf Design and Implementation of OpenXM} \\
 --- Open message eXchange protocol for Mathematics 
}
\author{ Masayuki Noro, 
         Nobuki Takayama\thanks{Department of Mathematics, Kobe University,\htmladdnormallink{http://www.math.kobe-u.ac.jp/$\tilde{\ }$taka}{http://www.math.kobe-u.ac.jp/\tilde{\ }taka}}}
\date{ November 17, 2000 }
*/
\maketitle

\def\noroa#1{  }
\def\remove#1{ }

//&C \noindent Draft for protocol version 1.1.3.  \\
//&C 1.1.3 is encoded as 001001003 in mathcap. \\

/*&jp
{\tt kxx/openxxx.tex}. {\bf Draft} 1997, 11/20 --- .
この文書は open XM の設計用のメモとしての役目もあるので,
一部のファイル名は開発者のみが参照できる.
*/

/*&C
\input{introduction}
\input{cmo-basic0}
\input{formal-expression}
\input{communication-model}
\input{stackmachine}
\input{project}
\input{control}
\input{start-session}
\input{string-expression}
\input{cmo-basic1}
\input{ox-localdata}
\input{error}
\input{redefine}
\input{implementation}
\input{howtowrite}
\input{cmo-register}
%\input{appendix}
\input{library}
*/

/*&C

\begin{thebibliography}{99}
\bibitem{GKW}
Gray, S., Kajler, N. and Wang, P. S.,
Design and Implementation of MP, a Protocol for Efficient
  Exchange of Mathematical Expressions,
  {\sl Journal of Symbolic Computation}, 1996.
\bibitem{gap}
Linton, S. and Solomon, A.,
OpenMath, IAMC and {\tt GAP},
preprint, 1999.
\bibitem{asir}
Noro, M. et al.,
A Computer Algebra System {\tt Risa/Asir},  1993, 1995, 2000\\
\htmladdnormallink{ftp://archives.cs.ehime-u.ac.jp/pub/asir2000/}{ftp://archives.cs.ehime-u.ac.jp/pub/asir2000/}
\bibitem{openmath}  \htmladdnormallink{\tt http://www.openmath.org}{http://www.openmath.org}
\bibitem{openxxx}   \htmladdnormallink{\tt http://www.math.kobe-u.ac.jp/OpenXM/}{http://www.math.kobe-u.ac.jp/OpenXM/}
\bibitem{openasir-intro}  Ohara, Takayama, Noro: Introduction to Open Asir , 
1999, (in Japanese),
Suushiki-Shyori, Vol 7, No 2, 2--17. (ISBN4-87243-086-7, SEG , Tokyo).
\bibitem{kan}
Takayama, N.,
{\em Kan: A system for computation in
algebraic analysis,} 1991 version 1,
1994 version 2, the latest version is 2.991106.
\htmladdnormallink{ftp://ftp.math.kobe-u.ac.jp/pub/kan}{ftp://ftp.math.kobe-u.ac.jp/pub/kan}
\bibitem{phc}
Verschelde, J.,
PHCpack: A general-purpose solver for polynomial systems by
homotopy continuation.  ACM Transaction on Mathematical Softwares, 25(2)
251-276, 1999.
\bibitem{iamc}
Wang, P.,
Design and Protocol for Internet Accessible Mathematical Computation.
Technical Report ICM-199901-001, ICM/Kent State University, 1999.
\bibitem{xml}
XML
\htmladdnormallink{\tt http://www.w3c.org}{http://www.w3c.org}
\end{thebibliography}
*/

/*&C
\bigbreak
\bigbreak
\bigbreak

\small{
\noindent
    \rightline{ Masayuki Noro,}
    \rightline{ FUJITSU LABORATORIES LTD., Kawasaki, Japan; (by Aug., 2000)}
 \rightline{{\tt noryo@flab.fujitsu.co.jp}}
   \rightline{Current Address: Department of Mathematics, Kobe University, 
               Rokko, Kobe, 657-8501, Japan;}
\rightline{{\tt noro@math.kobe-u.ac.jp}}

\vskip .5cm

\noindent
    \rightline{ Nobuki Takayama,}
   \rightline{Department of Mathematics, Kobe University, 
               Rokko, Kobe, 657-8501, Japan;}
\rightline{{\tt takayama@math.kobe-u.ac.jp}}
}

\end{document}
*/
